### 2017-12-11 manuscript.rmd
added Phil's changes

```
These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig. 3) [@Dunn:2005dy]. There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora. We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans. Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis. There is strong support for the Pyrostephidae as sister to all other non-Apolemid codonophorans. Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig. 3), which includes the remaining non-Apolemid, non-Pyrostephid "Physonectae". We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*. Within Euphysonectae, we identify Clade A and Clade B (Fig. 3). Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis]. This clade also shares the prescence of a descending mantle canal within the nectophores (Fig. S6). Clade B is a group of deep-sea siphonophores found within the bathypelagic. Within Clade B there is low support for the placement of *Erenna richardi*, and more taxon sampling will be required to determine the relationship of species within this clade.

Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae *sensu stricto*, a clade restricted to *Agalma*, the Athorybiidae, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic]. In the rDNA study, *Physophora hydrostatica* (sister species to *P. gilmeri*) was sister to the Forskaliidae with low support. The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp. is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae *sensu stricto*. Placement of *Cordagalma* outside Agalmatidae *sensu stricto* is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].

@@

One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character]. For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 10 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision]. Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none. Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome tree (Fig. 4).

@@

Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny). These patterns of loss are not captured by the present sampling.

Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore]. We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis]. We reconstruct them as present in the common ancestor of siphonophores (Fig. 5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans (with the exception of Stephanophyes [@totton1965synopsis]), and in *Marrus claudanielis*.

@@

## Conclusions
Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support. A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemid codonophorans. Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. Future studies with the addition of key taxa within the Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.

>>>>

```
These findings are largely consistent with a previous analysis based on two genes (16S and 18S ribosomal RNA) (Fig. 3) [@Dunn:2005dy]. There is strong support for the Cystonectae as sister to all other siphonophores, the Codonophora. We also find strong support for Calycophorae nested within the paraphyletic "Physonectae", and for the Apolemidae as sister to all other codonophorans. Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in this 1,071-gene transcriptome analysis. There is strong support for the Pyrostephidae as sister to all other non-apolemiid codonophorans. Within the clade that is sister to the Pyrostephidae, we find two main clades, the Calycophorae and a clade we call the Euphysonectae (Fig. 3), which includes the remaining non-apolemiid, non-pyrostephid "Physonectae". We define the Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*. Within Euphysonectae, we identify Clade A and Clade B (Fig. 3). Clade A is a group of siphonophores that all have an involucrum, or fold, that forms around the base of the tentilla [@totton1965synopsis]. This clade also shares the prescence of a descending mantle canal within the nectophores (Fig. S6), and is also monoecious (Fig. 5). Clade B is a group of deep-sea siphonophores found within the bathypelagic. Within Clade B there is low support for the placement of *Erenna richardi*, and more taxon sampling will be required to determine the relationship of species within this clade.

Within Clade A, *Physophora gilmeri* along with *Lychnagalma utricularia* (both not included in the previous phylogeny) are sister to the Agalmatidae, a clade restricted to *Agalma*, *Athorybia*,*Melophysa*, *Halistemma* and *Nanomia* [@Dunn:2005dy; @pugh2006taxonomic]. In the rDNA study, *Physophora hydrostatica* (sister species to *P. gilmeri*) was sister to the Forskaliidae with low support. The position of *Cordagalma cordiforme* (= *Cordagalma ordinatum*) [@pugh2016synopsis] was previously unresolved, while in this analysis *Cordagalma* sp. is in a clade with *Forskalia asymmetrica*, falling outside of the Agalmatidae. Placement of *Cordagalma* outside Agalmatidae is consistent with previous analyses using morphological and molecular data [@pugh2006taxonomic;@Dunn:2005dy].

@@

One of the most striking aspects of siphonophore biology is their diversity of unique zooid types [@Beklemishev1969;@cartwright2010character]. For example, the genus *Forskalia* has 5 basic zooid types (nectophore, gastrozooid, palpon, bract, and gonophore), and in some species, a total of 9 when considering zooid subtypes (4 types of bract, male & female gonophores)[@pugh2003revision]. Diphyomorphs have more than one type of propulsive nectophore, while cystonects have none. Here we reconstruct the evolutionary origins of several zooid types on the present transcriptome tree (Fig. 4).

@@

Bracts are highly reduced zooids unique to siphonophores, where they are only present in Codonophora (Fig. 4). Bracts are functional for protection of the delicate zooids and to help maintain neutral buoyancy [@jacobs1937beobachtungen]. Some calycophorans are able to actively exclude sulphate ions in their bracts to adjust their buoyancy along the colony [@bidigare1980role]. The common ancestor of Codonophora had only one bract subtype, which was lost in Hippopodiidae, *Physophora hydrostatica*, the Clausophyidae, and in *Gymnopraia lapislazula* (none of the latter three species and clades are included in the phylogeny). These patterns of loss are not captured by the present sampling.

Palpons are modified reduced gastrozooids used for digestion and circulation of the gastrovascular fluid [@mackie1987siphonophore]. We do not distinguish here between gonopalpons (palpons associated with gonodendra, without a tentacle, as in the cystonects) and palpons borne on the stem (typically with a reduced tentacle or palpacle) [@totton1965synopsis]. We reconstruct them as present in the common ancestor of siphonophores (Fig. 5B), retained in most species, but lost three times independently in the branches leading to *Pyrostephidae* (represented here by the genera *Bargmannia* and *Rudjakovia*), in calycophorans, and in *Marrus claudanielis*. Within the calycophorans, one species *Stephanophyes superba* (not included in this phylogeny) has polyp-like zooids that have been described as palpons [@totton1965synopsis], but the exact identity of this zooid is not clear and needs further morphological examination.

@@

## Conclusions
Using phylogenomic tools we were able to resolve deep relationships within the Siphonophora with strong support. A previous rDNA study was unable to resolve deep relationships within the Codonophora, while our study suggests that the Pyrostephidae are sister to all other non-apolemiid codonophorans. Our study supports the rejection of the monophyly of the Physonectae, a traditional morphological grouping, and rejects the monophyly of monoecious siphonophores. Future studies with the addition of key taxa within the Codonophora will enable resolution of internal relationships within Siphonophora, and shed light on patterns of character evolution that cannot be fully assessed here.
